---
title: "Stawberries"
author: MA615
date: 2024 Sept 25
format: pdf
editor: visual
---

# Preparing data for analysis

## Introduction: foundations

Before we begin to work with the strawberry data, let's talk about how we will approach the work.

### Data cleaning and organization

Cleaning and organizing data for analysis is an essential skill for data scientists. Serious data analyses must be presented with the data on which the results depend. The credibility of data analysis and modelling depends on the care taken in data preparation and organization.

#### References

In their handbook ["An introduction to data cleaning with R" by Edwin de Jonge and Mark van der Loo](https://cran.r-project.org/doc/contrib/de_Jonge+van_der_Loo-Introduction_to_data_cleaning_with_R.pdf), de Jonge and van der Loo go into detail about specific data cleaning isssues and how to handle them in R.

["Problems, Methods, and Challenges in Comprehensive Data Cleansing" by Heiko Müller and Johann-Christoph Freytag](https://www.researchgate.net/profile/Heiko-Mueller/publication/228929938_Problems_methods_and_challenges_in_comprehensive_data_cleansing/links/09e415101b58541e2c000000/Problems-methods-and-challenges-in-comprehensive-data-cleansing.pdf) is a good companion to the de Jonge and van der Loo handbook, offering additional issues in their discussion.

### Attitudes

Mechanistic descriptions of data cleaning methods are insufficient.

#### Data is the product (or by-product) of purposeful human activity

Much of the data used in analysis accessed on local databases or online which may create the impression that the data have been carefully curated. Beware. Data are produced by people for a purpose, with a point-of-view, and at a time and location that may affect the data. The provenance and lineage of the data are meta data you should include when reporting analysis. Data collection is purposeful human activity with all of the risks and weaknesses that are part of any purposeful human activity.

#### Data is language

Data has meaning. Data can be included in sentences related to the meaning of the data. Cleaning and organizing data should be informed by the meaning the data convey and how that meaning relates to the research you are doing do achieve this important result.

-   Immerse yourself in the data. Put data into context.

-   Visualize the data to find problems, confirm your understandings, and plan your data organization. People do a bad job of seeing meaningful patterns in data but a good job of seeing patterns of all kinds when data are rendered as plots. As you product and show visualizations, ask your self and those who view your presentations, "what do you see?" and "what do you wonder?"

## Example: Strawberries

### Public information

[WHO says strawberries may not be so safe for you--2017March16](https://med.news.am/eng/news/13621/who-says-strawberries-may-not-be-so-safe-for-you.html)

[Pesticides + poison gases = cheap, year-round strawberries 2019March20](https://www.ewg.org/foodnews/strawberries.php)

[Multistate Outbreak of Hepatitis A Virus Infections Linked to Fresh Organic Strawberries-2022March5](https://www.cdc.gov/hepatitis/outbreaks/fresh-strawberries-2022/?CDC_AAref_Val=https://www.cdc.gov/hepatitis/outbreaks/2022/hav-contaminated-food/index.htm)

[Strawberry makes list of cancer-fighting foods-2023May31](https://issuu.com/mechlocal/docs/053123_mech_asf/s/25386339)

## What is the question?

-   Where they are grown? By whom?

-   Are they really loaded with carcinogenic poisons?

-   Are they really good for your health? Bad for your health?

-   Are organic strawberries carriers of deadly diseases?

-   When I go to the Market should I buy conventional or organic strawberries?

## The data

The data set for this assignment has been selected from: \[[USDA_NASS_strawb_2024SEP25](https://quickstats.nass.usda.gov) <br> The data have been stored on NASS here: [USDA_NASS_strawb_2024SEP25](https://quickstats.nass.usda.gov/results/28141CE5-7429-36C3-A385-33ACC8030763)

and has been stored on the blackboard as strawberries25_v3.csv.

## USDA NASS

```{r}
#| label: load libraries
#| warning: false
#| message: false

library(knitr)  
library(kableExtra)
library(tidyverse)
library(stringr)
library(magrittr)
```

## Read the file

```{r}
#| label: read data - glimpse 

strawberry <- read_csv("strawberries25_v3.csv", col_names = TRUE)

#glimpse(strawberry)
```

Examine the data. How is it organized?

```{r}

## is every line associated with a state?

state_all <- strawberry |> distinct(State)

state_all1 <- strawberry |> group_by(State) |> count()

## every row is associated with a state

sum(state_all1$n) == dim(strawberry)[1]


## to get an idea of the data -- looking at california only

calif_census <- strawberry |> filter((State=="CALIFORNIA") & (Program=="CENSUS"))

calif_census <- calif_census |> select(Year, `Data Item`, Value)

### 

calif_survey <- strawberry |> filter((State=="CALIFORNIA") & (Program=="SURVEY"))

calif_survey <- strawberry |> select(Year, Period, `Data Item`, Value)

```

## Remove columns with a single value in all columns and county in `Geo Level`

```{r}
#| label: function def - drop 1-item columns
#| echo: false

drop_one_value_col <- function(df, prt_val = FALSE){ 
 # browser()
  df_id <- ensym(df)
  if(prt_val){
  msg = paste("Looking for single value columns in data frame: ",as.character(df_id) )
  print(msg)}
  ## takes whole dataframe
dropc <- NULL
val <- NULL
## test each column for a single value
for(i in 1:dim(df)[2]){   
  if(dim(distinct(df[,i]))[1] == 1){
    dropc <- c(dropc, i)
    val <- c(val, df[1,i])
  }
} 

if(prt_val){
if(is.null(dropc)){
  print("No columns dropped")
  return(df)}else{
   print("Columns dropped:")
   # print(colnames(df)[drop])
   print(unlist(val))
   df <- df[, -1*dropc]
   return(df)
  }
}
 df <- df[, -1*dropc]
   return(df)
}

## function shift_loc
## Moves adjacent data cells in a data.frame on a single row
## Use this function to fix alignment problems after separating 
## columns containing multiple columns of data. 

## Of course the working assumption is that there is room in the 
## data frame for the data you're shifting.
##
## The data cells that are empty after the data shift are NA.
## 
## Input paramaters
## 
## df -- data frame
## col_name -- name of colume where the left-most data item is located
## dat_name -- name of data item in the column
## num_col -- the number of columns is the same as the number of
##            adjacent data to be moved.
## num_shift -- the number of rows to move the data 
##


shift_loc <- function(df, col_name, dat_name, num_col, num_shift){
 # browser()
  col_num = which(colnames(df) == col_name)
  row_num = which(df[,col_num] == dat_name)  ## calcs a vector of rows
  
  for(k in 1:length(row_num)){
  d = rep(0,num_col) ## storage for items to be moved
  for(i in 1:num_col){
    d[i] = df[row_num[k], col_num + i - 1]
  }
  for(i in 1:num_col){
    ra = row_num[k]
    cb = col_num + i - 1
    df[ra, cb] <-  NA
  }
  for(j in 1:num_col){
    rc = row_num[k]
    cd = col_num + j - 1 + num_shift
    df[rc, cd] = d[j]
  }
  }
 # sprintf("Rows adjusted:")
  # print("%d",row_num)
  return(df)
}
```

```{r}

strawberry <- drop_one_value_col(strawberry)

drop_one_value_col(strawberry)

strawberry <- strawberry |>
  filter(`Geo Level` == "NATIONAL" | `Geo Level` == "STATE")
```

## Separate strawberry data set into small data sets to understand the data better

We separated `Census` and `Survey` data from the `strawberry` data set in order to examine the data better. Furthermore, `Data Item` column was separated into two column : `Fruit` and `Category`by `-` .

Focusing on the census data first, `Fruit` is further divided into `ORGANIC` and `Organic detail` leading us to get `organic` data set from census.

```{r}
#|label: split strawberries into census(further organic) and survey data
#|echo: false
census <- strawberry |> filter(Program == "CENSUS")

survey <- strawberry |> filter(Program == "SURVEY")
census <- census |> drop_one_value_col()

survey <- survey |> drop_one_value_col()

census <- census |>
  separate_wider_delim(  cols = `Data Item`,
                         delim = " - ",
                         names = c("Fruit",
                                 "Category"),
                         too_many = "error",
                         too_few = "align_start"
                       )
census <- census |>
  separate_wider_delim(  cols = Fruit,
                         delim = ", ",
                         names = c("Fruit",
                                 "ORGANIC",
                                 "Organic detail"),
                         
                         too_many = "error",
                         too_few = "align_start"
                       )

census <- census |> drop_one_value_col()
organic <- census |> filter(ORGANIC == "ORGANIC")
census <- census[(is.na(census$ORGANIC)),]
census <- census |> drop_one_value_col()
```

Split `Category` by `" "` into `Measure` and `Bearing Type` and consequently, removing `WITH` from `Bearing Type`.

```{r}
census <- census |>
  separate_wider_delim(  cols = `Category`,
                         delim = " ",
                         names = c("Measure",
                                 "Bearing Type"),
                         too_many = "merge",
                         too_few = "align_start"
                       )
  
census$`Bearing Type` <- str_replace(census$`Bearing Type`, "WITH ", "")
```

Upon observing `Domain Category` as per instruction of assignment 1 in strawberry, we just rename `Domain Category` into `size_bracket` for the `census` as it is majorly having size brackets for `Domain Category` Column. Along with it, `NOT SPECIFIED` is renamed into `TOTAL` and `AREA GROWN:` is removed. This cleans up the `census` data.

```{r}
census <- census |> rename(size_bracket = `Domain Category`)

census$size_bracket <- str_replace(census$size_bracket, "NOT SPECIFIED", "TOTAL")

census$size_bracket <- str_replace(census$size_bracket, "AREA GROWN: ", "")
organic <- organic |> drop_one_value_col()
```

Similarly to `census`, `Data Item` is split into four columns `Fruit`,`Category`,`Measure` and `Metric` for `survey` data by `,`. `Fruit` is further separated into `Fruit` and `Applications`

```{r}
survey <- survey |>  separate_wider_delim(cols = `Data Item`,
                                                delim = ", ",
                                                names = c("Fruit",
                                                         "Category",
                                                         "Measure",
                                                         "Metric"
                                                         ),
                                                too_many = "merge",
                                                too_few = "align_start")
                                                
                                                  
survey <- survey |> separate_wider_delim(cols = "Fruit", 
                                                 delim = " - ",
                                                 names = c("Fruit",
                                                           "Application"),
                                                 too_many = "merge",
                                                 too_few = "align_start"
                                                 )
  
```

#    Fixing Misplaced Values

Using `shift_loc`, some values that are supposed to be in other column (here column to the right of `Application`) are searched in the `Application` and then shifted right to the expressed number rows away .

```{r}
#|label: fix the misplaced values

survey %<>% shift_loc("Application", "PRICE RECEIVED", 2, 1 )

survey %<>% shift_loc("Application", "ACRES HARVESTED", 1, 1 )

survey %<>% shift_loc("Application", "ACRES PLANTED", 1, 1 )

survey %<>% shift_loc("Application", "PRODUCTION", 2, 1 )

survey %<>% shift_loc("Application", "YIELD", 2, 1 )

survey %<>% shift_loc("Application", "APPLICATIONS", 3, 1 )

survey %<>% shift_loc("Application", "TREATED", 3, 1 )

survey %<>% drop_one_value_col()

```

In `survey`, `Domain` is further separated into `Chemical` and `Type`. We then filter `TOTAL(survey_total),` `CHEMICAL(survey_chem)` and `FERTILIZER(survey_chem)` data sets from the `survey`

```{r}
survey <- survey |>  
  separate_wider_delim(cols = Domain,
                      delim = ", ",
                      names = c("Chemical",
                                "Type"),
                                                
                      too_many = "merge",
                      too_few = "align_start")
survey_total <- survey |>  filter(Chemical == "TOTAL")
survey_chem <- survey |>  filter(Chemical== "CHEMICAL")
survey_fert <- survey |>  filter(Chemical == "FERTILIZER")
```

Similar to the logic we applied at `Application` , we apply it on the `Measure` as well to fill in the NAs in the right places. Further, `Category` is divided into `Market` and `Action`

```{r}
survey_total %<>% drop_one_value_col()

### align terms

survey_total %<>% shift_loc("Measure", "MEASURED IN $ / CWT", 1, 1 )


survey_total %<>% shift_loc("Measure", "MEASURED IN $", 1, 1 )


survey_total %<>% shift_loc("Measure", "MEASURED IN CWT", 1, 1 )

survey_total %<>% shift_loc("Measure", "MEASURED IN TONS", 1, 1 )


survey_total %<>% shift_loc("Measure", "MEASURED IN CWT / ACRE", 1, 1 )

survey_total %<>% shift_loc("Measure", "MEASURED IN TONS / ACRE", 1, 1 )


survey_total <- survey_total |>  
  separate_wider_delim(cols = Category,
                      delim = " - ",
                      names = c("Market",
                                "Action"),
                   too_many = "merge",
                    too_few = "align_start")

```

Shifting values from `Market` to the right places. This cleans up `survey_total`

```{r}
survey_total %<>%
  select(-`State ANSI`)
survey_total <-  survey_total |>
  shift_loc("Market", "PRODUCTION", 2, 1)

survey_total <-  survey_total |>
  shift_loc("Market", "PRICE RECEIVED", 2, 1)
 
```

`Category` in `survey_chem` is divided into two categories namely `cat1` and `cat2`. Further due to repeating words, we remove `MEASURED IN` and `CHEMICAL` from the `Measure` and `Domain Category` respectively. We get `Chemical Name` from the `Domain Category` after seperating it into two and removing the first column. Punctuation signs are removed from the `Chemical Name` which we later divide into `Chemical Name` and `Code`. This cleans up `survey_chem`.

```{r}
survey_chem <- survey_chem |> drop_one_value_col()

survey_chem <- survey_chem |> select(-`State ANSI`)

survey_chem <- survey_chem |>  
  separate_wider_delim(cols = Category,
                      delim = " - ",
                      names = c("cat1",
                                "cat2"),
                   too_many = "merge",
                    too_few = "align_start")
survey_chem$Measure <- str_replace(survey_chem$Measure, "MEASURED IN ", "")

survey_chem$`Domain Category` <- str_replace(survey_chem$`Domain Category`, "CHEMICAL, ", "")

survey_chem <- survey_chem |>
          separate_wider_delim(cols = `Domain Category`,
                              delim = ": ",
                              names = c("type",
                                "Chemical Name"),
                           too_many = "merge",
                            too_few = "align_start")

survey_chem <- survey_chem |> select(-type)
  
survey_chem$`Chemical Name` <- str_replace(survey_chem$`Chemical Name`, "^\\(", "")
  
survey_chem$`Chemical Name` <- str_replace(survey_chem$`Chemical Name`, "\\)$", "")
  
survey_chem <- survey_chem |>  
  separate_wider_delim(cols = `Chemical Name`,
                      delim = " = ",
                      names = c("Chemical Name",
                                "Code"),
                   too_many = "error",
                    too_few = "align_start")

```

Now, we are to clean the `survey_fert` containing fertilizers data from the `survey`. Similarly to the `survey_chem`, we divide `Category` into two columns by `-` and then remove `MEASURED IN` and `CHEMICAL` from the `Domain Category.` Following the same routine, we clean up `survey_fert`.

```{r}
survey_fert <- survey_fert |> drop_one_value_col()

survey_fert <- survey_fert |> select(-`State ANSI`)

survey_fert <- survey_fert |>  
  separate_wider_delim(cols = Category,
                      delim = " - ",
                      names = c("cat1",
                                "cat2"),
                   too_many = "merge",
                    too_few = "align_start")
                
survey_fert$Measure <- str_replace(survey_fert$Measure, "MEASURED IN ", "")

survey_fert$`Domain Category` <- str_replace(survey_fert$`Domain Category`, "CHEMICAL, ", "")

survey_fert <- survey_fert |>
          separate_wider_delim(cols = `Domain Category`,
                              delim = ": ",
                              names = c("type",
                                "Chemical Name"),
                           too_many = "merge",
                            too_few = "align_start")
  
survey_fert$`Chemical Name` <- str_replace(survey_fert$`Chemical Name`, "^\\(", "")
  
survey_fert$`Chemical Name` <- str_replace(survey_fert$`Chemical Name`, "\\)$", "")
   
survey_fert <- survey_fert |> drop_one_value_col()

```

We convert `Value` column in `census` and `survey` into numeric.

```{r}
census$Value <- as.numeric(str_replace(census$Value, ",", ""))

organic$Value <- as.numeric(str_replace(organic$Value, ",", ""))
```

#  Imputation

We observe that in the `Bearing Type` we have some categories such as `BEARING`, `NON BEARING` and `GROWN.` Corresponding to the `TOTAL` value in the `Domain` and `size_bracket` , we can see the sum of all the previous value in the same category in `Value` Column. Using this logic, we get the function (with the help of chatgpt) `impute_values.` This will impute 0 where there are no previous values to apply the logic for `TOTAL` (these are very less cases). Similarly for more than one NA values in the same category. we get the remainder from subtracting sum of non-NA values from `TOTAL` to distribute remainder equally into NA values.

```{r}
library(dplyr)

impute_values <- function(df) {
  # Create a copy of the original data frame
  original_df <- df
  
  # Group by the relevant columns without changing the order
  df <- df %>%
    group_by(State, Measure, `Bearing Type`) %>%
    mutate(
      # Get the total value for the group
      total_value = Value[size_bracket == "TOTAL"],
      # Sum non-total values
      sum_non_total = sum(Value[size_bracket != "TOTAL"], na.rm = TRUE),
      # Count the number of NAs in the non-total values
      na_count = sum(is.na(Value[size_bracket != "TOTAL"])),
      # Fill NAs in non-total rows if there's a total value
      Value = ifelse(
        is.na(Value) & size_bracket != "TOTAL" & !is.na(total_value),
        round((total_value - sum_non_total) / na_count, 2),
        Value
      ),
      # Fill TOTAL if it is NA and non-total values are available
      Value = ifelse(
        size_bracket == "TOTAL" & is.na(Value),
        round(sum_non_total, 2),
        Value
      )
    ) %>%
    ungroup() %>%
    select(-total_value, -sum_non_total, -na_count) # Clean up intermediate columns

  # Format Value column to two decimal places
  original_df$Value <- round(df$Value, 2)
  
  return(original_df)
}

census <- impute_values(census)


```

There are only two states in this data set that are having chemical data i.e. `CALIFORNIA` and `FLORIDA`.We start by the seeing how many chemicals are there in the `survey_chem` (175).

```{r}

unique(survey_chem$State)

chemical_counts <- survey_chem %>%
  group_by(`Chemical Name`) %>%
  summarise(case_count = n()) %>%
  arrange(desc(case_count))

# View the result
print(chemical_counts)

```

### **Good Chemicals**

1.  **Neem Oil** (NEEM OIL, NEEM OIL, CLAR. HYD.)

2.  **Garlic Oil** (GARLIC OIL)

3.  **Canola Oil** (CANOLA OIL)

4.  **Sulfur** (SULFUR)

5.  **Bacillus Subtilis** (BACILLUS SUBTILIS)

6.  **Beauveria Bassiana** (BEAUVERIA BASSIANA)

7.  **Trichoderma Harzianum** (TRICHODERMA HARZ.)

8.  **Aureobasidium Pullulans** (AUREOBASIDIUM PULLULANS DSM 14940, AUREOBASIDIUM PULLULANS DSM 14941)

9.  **Hydrogen Peroxide** (HYDROGEN PEROXIDE)

10. **Mustard Oil** (MUSTARD OIL)

### **Bad Chemicals**

1.  **Glyphosate** (GLYPHOSATE ISO. SALT, GLYPHOSATE POT. SALT)

2.  **Malathion** (MALATHION)

3.  **Chlorpyrifos** (CHLORPYRIFOS)

4.  **Paraquat** (PARAQUAT)

5.  **Carbaryl** (CARBARYL)

6.  **Imidacloprid** (IMIDACLOPRID)

7.  **Bifenthrin** (BIFENTHRIN)

8.  **Permethrin** (PERMETHRIN)

9.  **Thiamethoxam** (THIAMETHOXAM)

10. **Mustard Oil** (MUSTARD OIL)

We check these oils as to which state are they in? Surprisingly they are in `CALIFORNIA`

```{r}
neem_oil_states <- survey_chem %>%
  filter(`Chemical Name` == "NEEM OIL" | `Chemical Name` == "NEEM OIL, CLAR. HYD." ) %>%
  select(State) %>%
  distinct()  # Get distinct states to avoid duplicates

neem_oil_states

garlic_oil_states <- survey_chem %>%
  filter(`Chemical Name` == "GARLIC OIL") %>%
  select(State) %>%
  distinct()  # Get distinct states to avoid duplicates

garlic_oil_states

```

This leads me to do further exploration on the number of cases per chemical for the both states. I see Oils used in `California` as one of the chemicals having `Major` cases i.e. 20

```{r}
# Assuming 'survey_chem' is your data frame and has a column for 'State' and 'Chemical Name'
california_chemicals <- survey_chem[survey_chem$State == "CALIFORNIA", ]

# Display the unique chemicals used in California
unique_california_chemicals <- unique(california_chemicals$`Chemical Name`)

# Count the occurrences of each chemical in California
california_counts <- table(california_chemicals$`Chemical Name`)

# Convert the table to a data frame
california_counts_df <- as.data.frame(california_counts)

# Rename the columns for clarity
colnames(california_counts_df) <- c("Chemical", "Count")
california_counts_df <- california_counts_df[order(-california_counts_df$Count), ]
row.names(california_counts_df) <- NULL
# Print the results
print(california_counts_df)

```

For the Florida, I observe that they are using Mustard oil in minority but still it is being used.

```{r}
# Assuming 'survey_chem' is your data frame and has a column for 'State' and 'Chemical Name'
florida_chemicals <- survey_chem[survey_chem$State == "FLORIDA", ]

# Display the unique chemicals used in California
unique_florida_chemicals <- unique(florida_chemicals$`Chemical Name`)

# Count the occurrences of each chemical in California
florida_counts <- table(florida_chemicals$`Chemical Name`)

# Convert the table to a data frame
florida_counts_df <- as.data.frame(florida_counts)

# Rename the columns for clarity
colnames(florida_counts_df) <- c("Chemical", "Count")
florida_counts_df <- florida_counts_df[order(-florida_counts_df$Count), ]
row.names(florida_counts_df) <- NULL
# Print the results
print(florida_counts_df)
```

I check how many cases for good chemicals (as in less toxicity or considered more beneficial than harmful) are there in both the states.

```{r}
good_chemicals <- c("NEEM OIL", "GARLIC OIL", "CANOLA OIL", "SULFUR", 
                    "BACILLUS SUBTILIS", "BEAUVERIA BASSIANA", 
                    "TRICHODERMA HARZ.", "AUREOBASIDIUM PULLULANS DSM 14940", 
                    "AUREOBASIDIUM PULLULANS DSM 14941", "HYDROGEN PEROXIDE", "MUSTARD OIL")

# Filter for good chemicals used in California
california_good_chemicals <- california_counts_df[california_counts_df$`Chemical` %in% good_chemicals, ]
row.names(california_good_chemicals) <- NULL
# Print the results
print(california_good_chemicals)
florida_good_chemicals <- florida_counts_df[florida_counts_df$`Chemical` %in% good_chemicals, ]
row.names(florida_good_chemicals) <- NULL
# Print the results
print(florida_good_chemicals)

```

For `CALIFORNIA`, I see out of 10 good chemicals chatgpt pointed out, there are 9 being used. Whereas for Florida, only 4 out of 10 are being used much to the disappointment but then again there are 175, I assume there would be more good chemicals that I am not observing being used in Florida. For the bad chemicals, both the states uses about 6 out of 10 I am observing.

```{r}
# Define bad chemicals
bad_chemicals <- c("BACILLUS THURINGIENSIS", "BIFENTHRIN", "CHLORPYRIFOS", 
                   "DIAZINON", "FIPRONIL", "IMIDACLOPRID", 
                   "MALATHION", "METOLACHLOR", "PERMETHRIN", 
                   "PENOXSULAM", "OXAMYL", "GLYPHOSATE")

# Filter for bad chemicals used in California
california_bad_chemicals <- california_counts_df[california_counts_df$`Chemical` %in% bad_chemicals, ]
row.names(california_bad_chemicals) <- NULL
# Print the results for California
print(california_bad_chemicals)

# Filter for bad chemicals used in Florida
florida_bad_chemicals <- florida_counts_df[florida_counts_df$`Chemical` %in% bad_chemicals, ]
row.names(florida_bad_chemicals) <- NULL
# Print the results for Florida
print(florida_bad_chemicals)

```

## Split Sales , Chemicals, Organic and Non-Organic into Different Dataframes

Writing Code into different CSV Files

```{r}
write.csv(strawberry, file = "strawberry_cleaneddata.csv")
write.csv(census, file = "census_data.csv")
write.csv(survey, file = "survey_data.csv")
write.csv(organic, file = "organic.csv")
#write.csv(census_non_organic, file = "census_non_organic.csv")
write.csv(survey_chem, file = "survey_chemical.csv")
write.csv(survey_total, file = "survey_total.csv")
write.csv(survey_fert, file = "survey_fert.csv")
```

# Questions after EDA

1.  Is there any connection between majority of Indian being in `California` and Oils being used as pesticides/fungicides especially Neem Oil?
2.  Why is Mustard Oil not used in the `California`?

# Study Material

[epa numbers](https://archive.epa.gov/pesticides/chemicalsearch/chemical/foia/web/html/128810.html)

[Active Pesticide Product Registration Informational Listing](https://ordspub.epa.gov/ords/pesticides/f?p=APPRIL_PUBLIC:2::::::)

[CAS for Methyl Bromide](https://ordspub.epa.gov/ords/pesticides/f?p=113:17::::::)

[pesticide chemical search](https://ordspub.epa.gov/ords/pesticides/f?p=chemicalsearch:1)

[toxic chemical dashboard](https://comptox.epa.gov/dashboard/)

[pubChem](https://pubchem.ncbi.nlm.nih.gov/)

The EPA PC (Pesticide Chemical) Code is a unique chemical code number assigned by the EPA to a particular pesticide active ingredient, inert ingredient or mixture of active ingredients.

Investigating toxic pesticides

[start here with chem PC code](https://ordspub.epa.gov/ords/pesticides/f?p=chemicalsearch:1)

[step 2](https://ordspub.epa.gov/ords/pesticides/f?p=113:1::::RP,17,1::) to get label (with warnings) for products using the chemical

[Pesticide Product and Label System](https://ordspub.epa.gov/ords/pesticides/f?p=113:1::::RP,17,1::)

[Search by Chemical](https://ordspub.epa.gov/ords/pesticides/f?p=113:17::::::)

[CompTox Chemicals Dashboard](https://comptox.epa.gov/dashboard/)

[Active Pesticide Product Registration Informational Listing](https://ordspub.epa.gov/ords/pesticides/f?p=APPRIL_PUBLIC:2::::::)

[OSHA chemical database](https://www.osha.gov/chemicaldata)

[Pesticide Ingredients](http://npic.orst.edu/ingred/)

[NPIC Product Research Online (NPRO)](http://npic.orst.edu/NPRO/)

[Databases for Chemical Information](http://npic.orst.edu/ingred/cheminfo.html)

[Pesticide Active Ingredients](http://npic.orst.edu/ingred/active.html)

[TSCA Chemical Substance Inventory](https://www.epa.gov/tsca-inventory)

[glyphosate](https://ordspub.epa.gov/ords/pesticides/f?p=CHEMICALSEARCH:3::::1,3,31,7,12,25:P3_XCHEMICAL_ID:2478)
